KIN: A method to infer relatedness from low-coverage ancient DNA

### 2022-09-07 paper_overleaf/main.tex
Updates from Overleaf

\paragraph{ROH detection:}
We first evaluated the performance of the ROH-HMM ( Fig.\ref{fig2:ROH}). We computed the proportion of ROH in all windows and compared it to the posterior probability of the ROH state, by computing the bias (i.e., the difference between both) and its variance across windows. Without ROH, we correctly inferred the absence of ROH as the observed bias ranges from $0.3$ to $0.7\%$ (variance $\leq0.006$; Table \ref{tab:Table 2}). With ROH, we observed a slight bias of $5$ to $9\%$ more ROH than simulated (variance $\leq0.004$). The largest bias observed ($9\%$) was for the datasets that included contamination. Ascertainment bias had a minimal effect on the detection of ROH (Table \ref{tab:Table 2}). Surprisingly, the bias was lower for lower coverages, perhaps because there is only one possible ROH state per window to model a continuous variable, viz. proportion of window in ROH, and this state is inferred more confidently with more data. Overall, the model predicts ROH well in all cases, even at a coverage of 0.1x (Fig.\ref{fig2:ROH}).

\begin{figure}[h!]
    \includegraphics[width=16cm]{plots/inkscape_finalImg/ROHplot_final1.png}
    \centering
    \caption{Estimation of ROH probabilities along the genome in simulations. Top panel shows proportion of differences in a simulated individual in windows along the genome, and the bottom panel shows probability of seeing no ROH. (A) Simulation with no ascertainment, contamination, or ROH. (B) Simulation with ROH. (C) Simulation with ROH and ascertainment. (D) Simulation with ROH and contamination.}
    \label{fig2:ROH}
\end{figure}

\begin{table}[h!]
\caption{\label{tab:Table 2}Model performance for ROH prediction. Here we test ROH-HMM in four different cases of simulations: Control (without ROH, ascertainment bias or contamination), R (with ROH), RA (with ROH and ascertainment bias), RC (with ROH and contamination)}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
    Coverage & Control(bias) & Control(var) & R(bias) & R(var) & RA(bias) & RA(var) & RC(bias) & RC(var)\\
    \hline
    4x & 0.007 & 0.006 & 0.081 & 0.036 & 0.083 & 0.036 & 0.091 & 0.039\\
    \hline
    0.2x & 0.003 & 0.003 & 0.068 & 0.03 & 0.068 & 0.026 & 0.076 & 0.029\\
    \hline
    0.1x & 0.003 & 0.002 & 0.049 & 0.021 & 0.05 & 0.021 & 0.057 & 0.022\\
    \hline
\end{tabular}
\label{table1}
\end{table}

>>>>

\paragraph{ROH detection:}
We first evaluated the performance of the ROH-HMM ( Fig.\ref{fig2:ROH}). We computed the proportion of ROH in all windows and compared it to the posterior probability of the ROH state, by computing the bias (i.e., the difference between both) and its variance across windows. Without ROH, we correctly inferred the absence of ROH as the observed bias ranges from $0.3$ to $0.7\%$ (variance $\leq0.006$; Table \ref{tab:Table 2}). With ROH, we observed a slight bias of $5$ to $9\%$ more ROH than simulated (variance $\leq0.04$). The largest bias observed ($10.1\%$) was for the datasets that included contamination. Ascertainment bias had a minimal effect on the detection of ROH (Table \ref{tab:Table 2}). Surprisingly, the bias was lower for lower coverages, perhaps because there is only one possible ROH state per window to model a continuous variable, viz. proportion of window in ROH, and this state is inferred more confidently with more data. Overall, the model predicts ROH well in all cases, even at a coverage of 0.1x (Fig.\ref{fig2:ROH}).

\begin{figure}[h!]
    \includegraphics[width=16cm]{plots/inkscape_finalImg/ROHplot_review.png}
    \centering
    \caption{Estimation of ROH probabilities along the genome in simulations. Top panel shows proportion of differences in a simulated individual in windows along the genome, and the bottom panel shows probability of seeing no ROH. (A) Simulation with no ascertainment, contamination, or ROH. (B) Simulation with ROH. (C) Simulation with ROH and ascertainment. (D) Simulation with ROH and contamination.}
    \label{fig2:ROH}
\end{figure}

\begin{table}[h!]
\caption{\label{tab:Table 2}Model performance for ROH prediction. Here we test ROH-HMM in four different cases of simulations: Control (without ROH, ascertainment bias or contamination), R (with ROH), RA (with ROH and ascertainment bias), RC (with ROH and contamination)}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
    Coverage & Control(bias) & Control(var) & R(bias) & R(var) & RA(bias) & RA(var) & RC(bias) & RC(var)\\
    \hline
    4x & 0.007 & 0.006 & 0.087 & 0.038 & 0.089 & 0.039 & 0.101 & 0.042\\
    \hline
    0.5x & 0.006 & 0.005 & 0.085 & 0.036 & 0.087 & 0.036 & 0.098 & 0.039\\
    \hline
    0.2x & 0.004 & 0.003 & 0.073 & 0.03 & 0.073 & 0.028 & 0.084 & 0.031\\
    \hline
    0.1x & 0.003 & 0.002 & 0.054 & 0.023 & 0.054 & 0.023 & 0.065 & 0.025\\
    \hline
\end{tabular}
\label{table1}
\end{table}
